---
title: "cluster"
author: "Jenny Li, Liz Cao, Kristy Ma"
date: '2022-04-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, message=FALSE, warning=FALSE)
```

```{r, library}
library(dplyr)
library(readr)
library(ggplot2)
library(tidymodels)
library(probably)
library(vip)
tidymodels_prefer()
theme_set(theme_bw())       
Sys.setlocale("LC_TIME", "English")
set.seed(74)
```

```{r, reading data}
breastCa<-read_csv(file = "breast-cancer.csv")
```
## Data Cleaning
```{r}
breastCa_Re<-breastCa %>% 
  drop_na() %>% 
  select(-c(13:22)) %>% 
  select(-1)

breastCa_Re_new<-breastCa_Re%>%
  mutate(concave_points_mean=`concave points_mean`)%>%
  select(-10) 
```

#Research goal
>With the clustering method, we plan to inspect the data and have the a better understanding of the following questions: What similarities are there among the tumors ?
Do there appear to be different kinds of tumors? If so, how many kinds are there ?

# Visual explorations

> For simplicity and visulizaiton purpose, we decided to first explore clustering based on the two most important variable on describing tumors:perimeter_mean and concave points_worst. These two chosen variable are mainly based on the variable importance measure from the classification method: perimeter_mean from LASSO with Logistic Regression and concave points_worst from random forest.

```{r}
ggplot(breastCa_Re_new, aes(x = perimeter_mean, y = `concave points_worst`)) +
    geom_point() +
    theme_classic()
```

#K-means clustering on perimeter_mean and concave points_worst

>We decide to run k-means for k=2 clusters based on the common sense that there are two diagnoses for a tumor: Malignant or Benign. We would ajust and find the optimal number of K if the result with k=2 is not satified.

```{r}
# Select just the perimeter_mean and concave points_worst variables
breastCa_Re_new_sub <- breastCa_Re_new %>%
    select(perimeter_mean, `concave points_worst`)

# Run k-means for k = centers = 2
set.seed(253)
kclust_k2 <- kmeans(breastCa_Re_new_sub, centers = 2)

# Display the cluster assignments
kclust_k2$cluster
```
```{r}
# Add a variable (kclust_k2) to the original dataset 
# containing the cluster assignments
breastCa_Re_new <- breastCa_Re_new %>%
    mutate(kclust_2 = factor(kclust_k2$cluster))
```

```{r}
# Visualize the cluster assignments on the original scatterplot
ggplot(breastCa_Re_new,
       aes(x = perimeter_mean, y = `concave points_worst`, color = kclust_2)) +
  geom_point() +
  theme_classic()
```
#Addressing variable scale

```{r}
# Run k-means on the *scaled* data (all variables have SD = 1)
set.seed(253)
kclust_k2_scale <- kmeans(scale(breastCa_Re_new_sub), centers = 2)
breastCa_Re_new <- breastCa_Re_new %>%
    mutate(kclust_2_scale = factor(kclust_k2_scale$cluster))

# Visualize the new cluster assignments
ggplot(breastCa_Re_new,
       aes(x = perimeter_mean, y = `concave points_worst`, color = kclust_2)) +
  geom_point() +
  theme_classic()
```

#Clustering on more variables
```{r}
# Select the variables to be used in clustering
breastCa_Re_new_sub2 <- breastCa_Re_new %>%
    select(c(2:21))

# Look at summary statistics of the 3 variables
summary(breastCa_Re_new_sub2)
```

```{r}
set.seed(253)
kclust_k2_allvars <- kmeans(scale(breastCa_Re_new_sub2), centers = 2)

breastCa_Re_new <- breastCa_Re_new %>%
    mutate(kclust_k2_allvars = factor(kclust_k2_allvars$cluster))


breastCa_Re_new %>%
  count(diagnosis,kclust_k2_allvars)
```

#Interpreting the clusters
```{r}
breastCa_Re_new %>%
    group_by(kclust_k2_allvars) %>%
    summarize(across(c(2:21), mean))
```

#Picking k
```{r}
# Data-specific function to cluster and calculate total within-cluster SS
breastCa_Re_new_cluster_ss <- function(k){
    # Perform clustering
    kclust <- kmeans(scale(breastCa_Re_new_sub2), centers = k)

    # Return the total within-cluster sum of squares
    return(kclust$tot.withinss)
}

tibble(
    k = 1:20,
    tot_wc_ss = purrr::map_dbl(1:20, breastCa_Re_new_cluster_ss)
) %>% 
    ggplot(aes(x = k, y = tot_wc_ss)) +
    geom_point() + 
    geom_line()+
    labs(x = "Number of clusters",y = 'Total within-cluster sum of squares') + 
    theme_classic()
```

#Normalized variables clustering visualizaiton of optimal K
```{r}
# Run k-means for k = centers = 3
set.seed(253)
kclust_k3 <- kmeans(breastCa_Re_new_sub, centers = 3)

# Display the cluster assignments
kclust_k3$cluster

# Run k-means on the *scaled* data (all variables have SD = 1)
set.seed(253)
kclust_k3_scale <- kmeans(scale(breastCa_Re_new_sub), centers = 3)
breastCa_Re_new <- breastCa_Re_new %>%
    mutate(kclust_3_scale = factor(kclust_k3_scale$cluster))

# Visualize the new cluster assignments
ggplot(breastCa_Re_new,
       aes(x = perimeter_mean, y = `concave points_worst`, color =kclust_3_scale)) +
  geom_point() +
  theme_classic()
```
# Perform clustering on more variabels with K=3
```{r}
set.seed(253)
kclust_k3_allvars <- kmeans(scale(breastCa_Re_new_sub2), centers = 3)

breastCa_Re_new <- breastCa_Re_new %>%
    mutate(kclust_k3_allvars = factor(kclust_k3_allvars$cluster))


breastCa_Re_new %>%
  count(diagnosis,kclust_k3_allvars)
```

#Interpreting the clusters with k=3
```{r}
breastCa_Re_new %>%
    group_by(kclust_k3_allvars) %>%
    summarize(across(c(2:21), mean))
```

